---
title: "Math185 HW6"
author: "Yuetong Lyu"
date: "5/28/2019"
output: pdf_document
---

Problem 1
The data set I chose has 25 people and each person has m gene expressions. I will divide them into two groups. The first 9 will be one group and the rest of them will be another. There are multiple genes in the data set also. Each gene is shared around these 2 groups.
```{r}
dat = read.table("GSE7621.csv", header = TRUE)
dat = dat[,-1]
m = nrow(dat)
pval = numeric(m)
for(i in 1:m){pval[i] = t.test(dat[i,1:9],dat[i,-(1:9)])$p.value}

reject = (pval <= 0.2)
R = sum(reject)
pval.bh = p.adjust(pval,"BH") #Assume data within samples are independent
pval.by = p.adjust(pval,"BY")

reject.bh = (pval.bh <= 0.2)
R.bh = sum(reject.bh)
reject.by = (pval.by <= 0.2)
R.by = sum(reject.by)

R
R.bh
R.by
```
After applying the FDR control at the level of 20%, we reduced the risk of having too many type 1 errors. By observing the result, there is a significant decrease in the rejections after we applied the FDR control. Since the "BH" method required data to be independent, we conclude that there are 3 genes that are differentially expressed.

Problem 2A
```{r}
load("dataset-chns-2006-subset3.Rdata")
diastolic = data$diastolic
systolic = data$systolic
plot(data$diastolic, data$systolic, pch = 16, , cex = 1, main = "Systolic vs Diastolic Values", xlab = "Diastolic Value", ylab = "Systolic Value")
```

Problem 2B
Linear regression 
H0: Beta1 = 0 H1: Beta1 != 0
```{r}
summary(lm(systolic ~ diastolic))
cor.test(diastolic, systolic) #Assume the data are normal
```
After observing the scatter plot, I decided to use the linear regression model to conduct a test. Given the p value is extremely small, we reject our null hypothesis and conclude that there is an association between Diastolic and Systolic. In addition, I checked by using correlation two sample test and also reached the same conclusion.

Problem 2C
```{r}
fit = summary(lm(systolic ~ diastolic))
fit$adj.r.squared
plot(data$diastolic, data$systolic, pch = 16, , cex = 1, main = "Systolic vs Diastolic Values", xlab = "Diastolic Value", ylab = "Systolic Value")
abline(lm(systolic ~ diastolic), col = "red", lwd = 2)
confint(lm(systolic ~ diastolic))
```
Assumption is error terms are iid N(0,sigma^2)

Problem 2D
```{r}
#fitting polynomial of degree 2
fit2 = lm(systolic ~ poly(diastolic, 2, raw = TRUE))
fit2R = summary(lm(systolic ~ poly(diastolic, 2, raw = TRUE)))$adj.r.squared

#40 and 180 comes from summary of diastolic 
pts = seq(40,180,len=1000)
val = predict(fit2, data.frame(diastolic = pts))
plot(data$diastolic, data$systolic, pch = 16, , cex = 1, main = "Systolic vs Diastolic Values", xlab = "Diastolic Value", ylab = "Systolic Value")
abline(lm(systolic ~ diastolic), col = "red", lwd = 2)
lines(pts, val, col = "blue", lwd = 2)
legend(130, 110, c("Degree 1", "Degree 2"), fill = c("red", "blue"))
```
Since there is no significant change in adjusted R squared, we choose to use linear model which is simpler than polynomial of degree 2.
