---
title: "Math185 HW1"
author: "Yuetong Lyu"
date: "4/13/2019"
output: pdf_document
---

Problem 1A
```{r}
chisq.power=function(k,t,n,B = 2000){
R <- numeric(B)
#build Pt
Pt <- numeric(2*k)
for(i in 1:k){Pt[i]=1/(2*k)+t}
for(i in (k+1):(2*k)){Pt[i]=1/(2*k)-t}
#simulate n elements from 1:2*k by prob Pt 
for(b in 1:B){
  ptsample <- sample(1:(2*k),n,replace = T, prob = Pt)
  tloop <- table(ptsample)
  if(chisq.test(tloop)$p.value <= 0.05){R[b]=1}
}
return(sum(R)/B)
}
```

Problem 1B
```{r}
k = 6
n = 500
t <- seq(1/(100*k), 1/(2*k)-1/(100*k),1/(100*k))
idx = 1
power <- numeric(length(t))
for(i in t){
  power[idx] <- chisq.power(k, i, n, B=2000)
  idx = idx + 1
}
plot(t, power, main="Power Curve when k = 6, n = 500", xlab = "t", ylab = "power")
```


Problem 2
```{r}
setwd("~/Desktop/math185_files")
Natality <- read.table("Natality.txt", header = T)
save(Natality, file = "natality-california-2017.rda")
load(file = "natality-california-2017.rda")
data <- matrix(NA,2,36)
for(i in 1:36){data[1,i] <- Natality$Births[i]}
for(i in 37:72){data[2,i-36] <- Natality$Births[i]}
rownames(data) <- c("Female", "Male")
colnames(data) <- Natality$County[1:36]
data
barplot(data,beside = T,legend = T,args.legend = list(x = "topright"))
chisq.test(data)
```
Since the p-value is 0.215, it is usually not small enough for us to reject the null hypothesis. Thus, we should keep the null hypothesis and conclude that gender and county are two independent elements, which means the chance of a girl being born does not depend on the county she is at. 


Problem 3
```{r}
chisq.perm.test = function(tb, B = 2000){
  
  #observed statistic
  Dobserve <- chisq.test(tb)$stat
  
  #now calculate useful numbers from tab
  numrow <- nrow(tb)
  numcol <- ncol(tb)
  totalcounts <- sum(tb[,])
  
  #idea: number each of the elements of the categories in the rows and columns 
  #how many numbers are in each row, it is a vector
  rows <- numeric(numrow)
  vecrow <- numeric(0)
  for (i in 1:numrow){
    rows[i] <- sum(tb[i,])
    vecrow <- c(vecrow, rep.int(i,times = rows[i]))
  }
  
  #how many numbers are in each column, it is a vector
  columns <- numeric(numcol)
  veccol <- numeric(0)
  for(j in 1:numcol){
    columns[j] <- sum(tb[,j])
    veccol <- c(veccol, rep.int(j, times = columns[j]))
    }
  #num is amount of times the statistcs is larger than Dobserved
  num = 0
  D <- numeric(B)
  
  for(b in 1:B){
    xvalues <- vecrow
    yvalues <- sample(veccol, totalcounts, replace = F) #permutation
    matvalues <- cbind(xvalues, yvalues) #n*2 matrix
    
    #restore the generated vector into a bootstrap matrix
    matnew <- matrix(0, nrow = numrow, ncol = numcol)
    
    #most complicated part
    for(indx in 1:numrow){
      for(indy in 1:numcol){
        for(indm in 1:totalcounts){
          if(all(matvalues[indm,] == c(indx, indy)) == T){
            matnew[indx,indy] <- matnew[indx,indy]+1
          }
        }
      }
    }
    D[b] <- chisq.test(matnew)$stat
    if(D[b]>=Dobserve){num = num + 1}
  }
  return((num+1)/(B+1))
}
HairEyeColor
temp <- apply(HairEyeColor,c(1,2),sum)
colnames(temp) = NULL
rownames(temp) = NULL
chisq.perm.test(temp, B = 2000)
chisq.test(temp, sim = T) #check by plug-in function
```


Problem 4A
```{r}
setwd("~/Desktop/math185_files")
grants <- read.csv("grants.csv", header = T,sep = ",", fill = T)
#only select third col(state) and sixth col(model)
smtbl <- grants[,c(3,6)]
#remove RI rows
smtbl <- smtbl[-which(smtbl$State == "RI"), ]
save(smtbl, file = "school-improvement-2010.rda")
load(file = "school-improvement-2010.rda")
#turn the table info into joint counts table
jointcounts <- with(smtbl, table(State, Model.Selected))
jointcounts <- jointcounts[-which(rownames(jointcounts) == "RI"),-c(1)]
jointcounts
barplot(t(jointcounts),legend = T, args.legend = list(x="topright"))
chisq.test(jointcounts)
```
Given the p-value is very small, we reject the null hypothesis and conclude that there is an association between the model that each school selected and the state where each school was located.

Problem 4B
```{r warning = FALSE}
options(warn = -1)
chisq.perm.test(jointcounts, B = 2000)
chisq.test(jointcounts,sim = T) #double check by using plug-in function
```
In this case, the function from problem 3 is applicable and the results are both approximately zeros.