---
title: "Math185 HW3"
author: "Yuetong Lyu"
date: "4/28/2019"
output: pdf_document
---

Problem 1
```{r}
require(MASS)
options(warn=-1)
# to run back warning use: options(warn=0)

chisq.normal.test = function(x, B = 2000){
  n=length(x)
  #calculating MLE
  mle = fitdistr(x, "normal")
  mu = mle$est[1]
  sigma = mle$est[2]
  
  #histogram for outomatically selecting the bins
  H = hist(x, 30, col="grey", freq = FALSE)
  
  #expected counts for these particular bins, and original statistics
  prob = pnorm(H$breaks, mean=mu, sd=sigma)
  exp.counts = n*diff(prob)
  obs.counts = table(cut(x, H$breaks))
  test = chisq.test(obs.counts, p=exp.counts, rescale.p=TRUE) 
  D = as.numeric(test$stat) #observed value of the statistics
  
  D.boot = numeric(B)
  count = 0
  for (b in 1:B) {
    dat.boot = rnorm(n, mean = mu, sd = sigma) 
    H = hist(dat.boot, 30, plot = FALSE) 
    mu.boot = mean(dat.boot)
    sigma.boot = sd(dat.boot)
    prob = pnorm(H$breaks, mean = mu.boot, sd = sigma.boot) 
    exp.counts = n*diff(prob)
    obs.counts = table(cut(dat.boot, H$breaks))
    result = chisq.test(obs.counts, p=exp.counts, rescale.p = TRUE)
    D.boot[b] = as.numeric(result$stat) #observed value of the statistics
    if(D.boot[b]>=D) {count=count+1}
  }
  return ((count + 1)/(B+1))
}
```

Problem 2A
```{r}
setwd("~/Desktop/math185_files")
library(readxl)
S1data = read_excel("S1Dataset.xlsx", sheet=2) 
brazil = S1data[which(S1data$Country==1),] 
portugal = S1data[which(S1data$Country==2),] 
mozambique = S1data[which(S1data$Country==3),] 

brazilweight = brazil$Weight
portugalweight=portugal$Weight 
mozambiqueweight=mozambique$Weight

brazilweight = as.numeric(na.omit(brazilweight))
portugalweight= as.numeric(na.omit(portugalweight)) 
mozambiqueweight= as.numeric(na.omit(mozambiqueweight))
```

Histogram and overlay the normal distribution
```{r}
require("MASS")

#comparing histogram with fitted density
#brazilweight
fit = fitdistr(brazilweight, "normal")
mu = fit$est[1] 
sigma = fit$est[2]
H = hist(brazilweight, 30, col="grey", freq=FALSE, main="Histogram of Brazil Weight Overlayed by Normal Density", xlab = "weight") 
t = seq(33, 135, len = 1000)
lines(t, dnorm(t, mean = mu, sd = sigma), col=2, lwd = 6)

#portugalweight
fit = fitdistr(portugalweight, "normal")
mu = fit$est[1]
sigma = fit$est[2]
H = hist(portugalweight, 30, col="grey", freq=FALSE, main="Histogram of Portugal Weight Overlayed by Normal Density")
t = seq(33, 135, len = 1000)
lines(t, dnorm(t, mean=mu, sd=sigma), col=2, lwd = 6)

#mozambiqueweight
fit = fitdistr(mozambiqueweight, "normal")
mu = fit$est[1]
sigma = fit$est[2]
H = hist(mozambiqueweight, 30, col="grey", freq=FALSE, main="Histogram of Mozambique Weight Overlayed by Normal Density")
t = seq(33, 135, len = 1000)
lines(t, dnorm(t, mean=mu, sd=sigma), col=2, lwd = 6)
```

Problem 2B
```{r}
#test for normality using function in problem 1
chisq.normal.test(brazilweight,B = 2000)
chisq.normal.test(portugalweight, B = 2000)
chisq.normal.test(mozambiqueweight, B = 2000)
```

Problem 3
```{r}
brazil = S1data[which(S1data$Country==1),] 
portugal = S1data[which(S1data$Country==2),] 
mozambique = S1data[which(S1data$Country==3),] 

brazilweight = brazil$Weight
portugalweight=portugal$Weight 
mozambiqueweight=mozambique$Weight

brazilweight = as.numeric(na.omit(brazilweight))
portugalweight= as.numeric(na.omit(portugalweight)) 
mozambiqueweight= as.numeric(na.omit(mozambiqueweight))
```

Permutation Test Function of H0:X~Y  H1:X stochastically dominates Y
```{r}
pertest = function(x,y,B){
  nx = length(x)
  ny = length(y)
  testorig = mean(x) - mean(y)
  z = c(x, y)
  n = length(z)
  
  count = 0
  for(b in 1:B){
    temp = sample(z, n ,replace = FALSE)
    xtemp = temp[1:nx]
    ytemp = temp[(nx+1):n]
    meantemp = mean(xtemp) - mean(ytemp)
    if(meantemp >= testorig){count = count +1}
  }
  return((count + 1)/ (B + 1))
}

pertest(brazilweight, portugalweight, 10000)
pertest(brazilweight, mozambiqueweight, 10000)
pertest(portugalweight,mozambiqueweight, 10000)
```
Kolmogorov-Smimov Two-Sample Test HO: F(x)=F(y)  H1:the cdf of x lies above that of y
Since the test statistics follows Kolmogorov-Smimov distribution, p-value is the probability that the sample test statistic is larger than the test statistics under the Kolmogorov-Sirov distribution.
```{r}
ks.test(brazilweight, portugalweight, alternative = "greater")
ks.test(brazilweight, mozambiqueweight, alternative = "greater")
ks.test(portugalweight, mozambiqueweight, alternative = "greater")
```

Wilcoxen Rank Sum Test/Mann Whitney U Test HO:mu(x)=mu(y)  H1:mu(x) > mu(y) 
Since the test statistics approximately follows the normal distribution, p-value is the probability that the sample test statistics is larger than that under the normal distribution.
```{r}
wilcox.test(brazilweight, portugalweight, alternative = "greater")
wilcox.test(brazilweight, mozambiqueweight, alternative = "greater")
wilcox.test(mozambiqueweight, portugalweight, alternative = "greater")
```


